---
title: "Revised Blog Analysis"
author: "Brian Perron"
date: "September 12, 2014"
output: html_document
---

Standard initialization


```{r initialization, warning=FALSE, message=FALSE, echo=FALSE}
rm(list=ls())
libs <- c("tm", "tau", "stringr", "reshape2", "stringi",, "microbenchmark")
lapply(libs, require, character.only = TRUE)
```

Read the blog files and check the encoding
```{r readData, cache=TRUE}
blogs <- readLines("~/Git/capstoneCoursera/en_US/blogs.txt")
unique(stri_enc_mark(blogs))
```

File is way too big -- let's start with taking a random subsample (XX).
```{r reduceFile}
set.seed(33345)
sampleLines <- c(1:length(blogs))
sampleSize <- sample(sampleLines, length(blogs), replace = FALSE)
blogs <- blogs[sampleSize]
```

Do general cleaning of blogs
```{r gsubCleaning, cache=TRUE}
blogs <- tolower(blogs)
blogs <- gsub("/",   "", blogs)
blogs <- gsub("'", '', blogs)
blogs <- gsub("’", '', blogs)
blogs <- gsub("‘", '', blogs)
blogs <- gsub("′", '', blogs)
blogs <- gsub("”", '', blogs)
blogs <- gsub("“", '', blogs)
blogs <- gsub("@",   "", blogs)
blogs <- gsub("–", "", blogs)
blogs <- gsub("\\|", "", blogs)
blogs <- gsub("[!?,.]+", ".", blogs)
blogs <- gsub("…", "", blogs)
blogs <- gsub('[])(;:#%$^\\~{}[&+=@/"<>_]+', "", blogs)
blogs <- removeWords(blogs, stopwords("english"))
blogs <- stripWhitespace(blogs)
```




Create initial set of bi-grams and tri-grams
```{r cache=TRUE}
bigrams <- textcnt(blogs, n = 2, method="string")
bigrams <- bigrams[order(bigrams, decreasing = TRUE)]
#print(bigrams[1:10])
#print(trigrams[1:10])
```

Create bi-grams dataframe
```{r}
# Convert to data frame
bigrams.df <- as.data.frame(bigrams)

# Create a new variable containing bigrams (currently as row names)
bigrams.df$ng.1 <- rownames(bigrams.df)

# Split the new variable into individual tokens
bigrams.split <- colsplit(bigrams.df$ng.1, " ", c("ng1", "ng2"))

# Create a new data frame with the tokens and the bi-gram count
bigram.df.updated <- data.frame(bigrams.split, bigrams.df$bigrams)

names(bigram.df.updated)[3] <- "count"

bigram.df <- bigram.df.updated
bigram.df <- bigram.df[complete.cases(bigram.df),]
```

Lots of stuff created.  Let's clean it all up and just keep the `bigram.df` object.  We can add more to the concatenating object if needed. 

```{r cleanUp}
# Remove everything but X items
rm(list = setdiff(ls(), c("bigram.df")))
```

Now, we can write a simple function that uses the basic bigram model.  The user must enclose the word with quotation marks within the function.  

`Question:` If the user doesn't include quotes, can I automatically paste them into the function?

```{r bigramFunction}
bigram.f <- function(x){
     a <- match(x, bigram.df[,1]) # Match patterns - fix to exact word
     bigram.a <- bigram.df[a,] # Subset data
     bigram.a.ordered <- bigram.a[with(bigram.a, order(-count)),]
     bigram.final.5 <- bigram.a.ordered[1:5,]
     bigram.final.5
}
```

```{r examples}
bigram.f("beer")
bigram.f("food")
bigram.f("jump")
bigram.f("hello")


